### 2017-05-18 paper/documents/0.abstract.tex
Add files via upload

\begin{abstract}

This is abstract\cite{eggert2015classification}

\end{abstract}