### 2017-05-18 paper/documents/5.conclusion.tex

\section{CONCLUSION}
\label{sec:conclusion}

This is conclusion