### 2017-05-20 paper/documents/1.introduction.tex
- change manuscript structure

\section{Introduction}

This is Introduction also the related work.