### 2017-05-22 paper/documents/5.conclusion.tex

\section{Conclusion and Future Work}
\label{sec:conclusion}

This is conclusion