### 2017-05-22 paper/documents/2.background.tex

\section{Background}
\label{sec:background}

Dota2 is team based game in Multi-player Online Battle Arena (MOBA). Each team have five players. Every player will choose their hero in pre-match games. Currently, there is 113 heroes, 166 current items. Dota2 game design is separated the heroes with three types. The first is agility heroes type, the second is strength heroes type and the last is intelligent heroes type. Every heroes have own specialty and weakness. The items can be combine into upgrade-able item which is will give more status in heroes abilities.
\\
Hero selection is important aspect of the game. Team needs to balance the heroes with different abilities that able to fulfill certain of roles and strategies. In the beginning of the game, each player will get 650 gold to buy starting items. 
\\
Dota2 map layout split into three lanes. Each lanes has three defensive towers.