### 2017-06-02 paper/documents/2.background.tex

\section{Background}
\label{sec:background}

\textit{Dota 2} is team based game in Multi-player Online Battle Arena (MOBA). It is divided by two areas, the Dire and the Radiant. Each team have five players. Every player will choose their hero in pre-match games. Currently, there is 113 heroes, 166 current items. Dota2 game design is separated the heroes with three types. The first is agility heroes type, the second is strength heroes type and the last is intelligent heroes type. Every heroes have own specialty and weakness. The items can be combine into upgrade-able item which is will give more status in heroes abilities.

Hero selection is important aspect of the game. Team needs to balance the heroes with different abilities that able to fulfill certain of roles and strategies. In the beginning of the game, each player will get 650 gold to buy starting items.

Dota 2 map layout split into three lanes. Each lanes has three defensive towers. the Dire lanes and the Radiant lanes are quite different. the top of Radiant it is called off-lane , meanwhile in Dire, the top lane is called safe-lane. The bottom lane of Radiant is called safe-lane, while in Dire it is called off-lane. And the middle lane will remain same.

\subsection{Player Roles}
\label{sec:player_roles}

Player roles or play styles is the way how player will choose and try to act as their roles or styles in the games. It is important to note that these roles describe a different facet of play than classic player type classifications (e.g. after battle