### 2017-06-03 paper/documents/1.introduction.tex

\section{Introduction}

Computer games is a big business. Competitive gaming growing the interest in this business. The competitive gaming events, the so-called electronic sports (sPorts). State-of-the-art multi-player online batle arena (MOBA) games are the most popular and succesful games in this regard. Game tournaments always give significant awards with big prize of money. Mostly, professional player can make a living from the prize.

In this paper we analyzing the player roles using supervised machine learning (ML) , investigate the applicability of feature normalization to enhance accuracy of prediction results. We also comparing our result with previous paper. Christoph Eggert on his paper said that their most stable and performing classifier is logistic regression. We will use the same classification method with their work. The different is we will using ROpenDota to built the data set.

An approach to ML in computer games in general was proposed by Drachen et al.\cite{drachen2014skill} . They suggest using unsupervised learning algorithms, specifically k-means and \textit{Simplex Volume Maximization}, to cluster player behavior data.

Our investigation about applicability and performance of \textit{supervised machine learning} (ML) to classify player roles based on behavior in \textit{Dota 2}, the most popular MOBA game. Such information could be useful for game designer to better understand how their game design influences emergent gameplay and player behavior but also for players, both casual and professional, who want to analyze their own performance or who want to learn from the others. It could also support casters and moderators in commentating and presenting matches. Furthermore, this research might hold implications for social and other research concerned with (human behavior in) games. While ML has been applied to games and traditional sports, most works are either interested in questions like spatial behavior, trying to predict the match outcome, or otherwise trying to correlate performance to certain events or behaviors. In contrast, we aim at building a classifier that is largely independent of individual players performance and that is also not tied to the overall match outcome but that is able to identify a player's role in term of the non-formally defined roles established as common grounds within the \textit{Dota 2} or MOBA community.

This paper contributes to the state of the art in several ways: We provide an in-depth discussion and novel approaches regarding the construction of complex attributes from low level data extracted from Dota 2 replay files, together with an evaluation of these attributes with respect to different classifiers. Based on the resulting reduced set of attributes, we compare and discuss the performance of a range of supervised classification algorithms, including logistic regression, random forest decision trees, support vector machines (in combination with Sequential Minimal Optimization), naive Bayes and Bayesian networks, classifying both with a newly established larger set of player roles, as well as with a reduced set inspired by related work ~\cite{eggert2015classification}.