### 2017-06-05 paper/documents/0.abstract.tex
- using 12pt font size

\begin{abstract}

The state of the art, computer games are big business, which is also reflected in the growing interest in competitive gaming. \textit{Multi-player online battle arena} games are the most successful competitive gaming in computer games industry. Every players have own specific roles in this team based game. \textit{Dota 2} is the most popular for this games. This paper analyzing player roles and their behavior using \textit{supervised machine learning} and also investigates the applicability of feature normalization to enhance accuracy of prediction result.We provide an in-depth discussion and novel approaches for constructing complex attributes from low-level data extracted from replay files. Using attribute evaluation techniques, we are able to reduce a larger set of candidate attributes down to a manageable number. \textit{Random Forest} the decision trees algorithm is the most stable and best performing classifier in our results. Again, \textit{Random Forest} although not the best performing classifier  but it is proved to be very stable and well suited to this domain.

\end{abstract}