### 2017-06-06 paper/documents/2.background.tex

\section{Background}
\label{sec:background}

\textit{Dota 2} is team based game in \textit{Multi-player Online Battle Arena} (MOBA). It is divided by two areas, the \textit{Dire} and the \textit{Radiant}. Each team has five players. Every player will choose their hero in pre-match games. Currently, there is 113 heroes, 166 current items. \textit{Dota 2} game design has separated the heroes with three types. The first is agility heroes type, the second is strength heroes type and the last is intelligent heroes type. Every hero has own specialty and weakness. The items can be combined into a upgrade-able item which is will give more status in heroes abilities.

Hero selection is an important aspect of the game. The team needs to balance the heroes with different abilities that able to fulfill certain of roles and strategies. In the beginning of the game, each player will get 650 gold to buy starting items.

\textit{Dota 2} map layout split into three lanes. Each lanes has three defensive towers (see figure \ref{fig:map_lane}). the Dire lanes and the Radiant lanes are quite different. the top of Radiant it is called off-lane, meanwhile, in Dire, the top lane is called safe-lane. The bottom lane of Radiant is called safe-lane, while in Dire it is called off-lane. And the middle lane will remain same.

\subsection{Player Roles}
\label{sec:player_roles}

Player roles or play styles is the way how the player will choose and try to act as their roles or styles in the games. It is important to note that these roles describe a different facet of play than classic player type classifications (e.g. after Bartle \cite{bartle1996hearts}) which aim to classify expressed character traits of the players and were designed to match role-play style games. Comparison of videos, Online guides, and commentary of professional players and commentators are our sources for selection and characterization of the player roles. Player roles definitions and naming conventions will, in any case, differ slightly among the player base and shift over time as the game evolves, constituting another challenge for \textit{Machine Learning} applications in this area. We isolated five player roles for the main \textit{Machine Learning} task, which strike a balance between covering common play styles in great detail while leaving out some exotic styles which are rarely observed or are minor variations of the other styles. The isolated roles were: \textit{Carry} - who are usually weak and need protection early on but are very strong in later stages, often deciding games. Carry typically end up with a high amount of last hits, gold per minute and overall kills, but they can get them in quite different ways. \textit{Mid Laner} - this kind of playing style mostly will get middle lane position since early game. focusing on collecting more gold to obtain high percentage lane of gold in early ten minutes. \textit{Offlaner} - safe lane as their common lane. \textit{Dire} jungle is the safe lane for the Dire team, otherwise, the \textit{Radiant} jungle is the safe lane for the Radiant team. Usually, have a balance between \textit{networth} and efficiently lane gold in early ten minutes. \textit{Roaming Support} - support heroes mostly weak, roaming support will have roaming attribute from replay file, this player role will buy support item such as: \textit{Observer ward} and \textit{Sentry ward} for vision. \textit{Hard Support} - sometimes called the babysitter, because this playing style needs to do everything for supporting the Carry. Mostly, buy support items much more than roaming support.